Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-18 chi2012.tex
discussion

\subsection{Experimental Setup}
The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen by default.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

Twenty-six (26) participants responded, from which we selected the first 20.  Ten participants were female.  All of the participants spoke English, but 12 of them were non-native English speakers.  Seven were non-students; out of the remaining 13, eight studied computer science and 5 studied other subjects.   On average, each study took 40 mins to complete.

@@

When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they understood it \emph{very strongly} (5).  Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as useful, with one person being neutral on each.  On the statement ``DP is confusing'',  11 strongly disagreed, 6 disagreed, and 3 agreed.  Revisiting the time trials, we discovered that the 3 participants who rated DP as strongly confusing completed the trials statistically significantly slower than the rest of the group. Survey results were visualised in Figure \ref{fig:survey}.

@@

\section{Discussion}

In this section, we first re-visit our study hypotheses introduced in Methodology, using observations from results to support views on each.  We follow this with a discussion of limitations, concerning the design of the interface, the state of the prototype, and the scope and execution of the study, including potential confounding factors, and how this impacts interpretability of results. Finally, we discuss our current and follow-on plans for continuing this line of research.

\subsection{Revisiting the hypotheses: Does DesignPalette facilitate serendipitous data integration?}

This study tested how

@@

% DAN's  qual analysis of processes

In order to complete the tasks, the users followed a number of different processes. The most common process followed
was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
of these next to the set of choices. They then made their choice based on the value of this tally.

>>>>

\subsection{Experimental Setup}
The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen by default.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

Twenty-six (26) participants responded, from which we selected the first 20.  Ten participants were female.  All of the participants spoke English, but 12 of them were non-native English speakers.  Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.   On average, each study took 40 mins to complete.

@@

When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they agreed \emph{very strongly}.  Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.  On the statement ``DP is confusing'',  11 strongly disagreed, 6 disagreed, and 3 agreed.  Revisiting the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials statistically significantly slower than the rest of the group. Survey results were visualised in Figure \ref{fig:survey}.

@@

\section{Discussion and Limitations}

In this section, we first re-visit our study hypotheses introduced in Methodology, using observations from results to support views on each.  We follow this with a discussion of limitations regarding the design of the interface, the state of the prototype, and the design and execution of the study.  Finally, we discuss our current and follow-on plans for continuing this line of research.

\subsection{Revisiting the hypotheses: Does DataPalette facilitate serendipitous data integration?}

We set out to answer three different questions with the DataPalette study. The first concerned whether people understood and could use the system (h1), whether the system enabled people to mix and integrate heterogeneous data (h2), and whether the interface facilitated task completion (h3).  Using real data obtained from typical data sources identified during our prestudy (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.

We find substantial support for h1. All participants reported that they understood the tool, and that it was easy to use.   All participants managed to use the interface to view, organize and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.  More than half of the participants used DP's visualisation features (charting or map -- often several times during the trial -- suggesting that these features were useful and usable as well.

More convincing of participants' understanding of the system was their feedback, specifically feature requests and desired capabilities.  Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.  Perhaps the most interesting suggestion was also the most common -- the ability to selectively view the provenance of information after instances have been combined.  For example, after his trial, P15 said:

\begin{quote}
Properties are a tricky issue. Sometimes you don’t care [which data source] a property is from, like ``address'' or ``phone number'' - for these the way it’s done now is fine. But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you’re talking about ``Yelp rating'' or some random reviwer's rating.  You also don’t know what typical ratings are, whether 5 stars are much better than 4, etc.
\end{quote}

Pertaining to h2, all participants were able to work with multiple sets of heterogeneous data effectively.  With respect to integration specifically, a majority (80\%) of the particiapnts successfully and deliberately integrated data using drag-and-drop SameAs capabilities.  While single-paths were used, multi-path features were not as widely used, only by 4 participants. We believe that participants may not have realized that creating multiple path selections per group was possible, instead assuming it to be similar to the single ``Current path'' state per window of most file managers.

Ascertaining DP's impact on task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.  While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.  However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.  We also found that particiapnts justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.  Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.  Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly suggests that DP made the task easier to complete.

@@

% DAN's  qual analysis of processes
In order to complete the tasks, the users followed a number of different processes. The most common process followed
was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
of these next to the set of choices. They then made their choice based on the value of this tally.
